---
title: "snRNA-seq analysis progress - 1; Tabeta et al."
author: "JS KIM"
date: "2024-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Overview
Progress to analyze the single nucleus RNA-seq (snRNA-seq) of Arabidopsis 8-d-old cotyledons of two genotypes, wild-type and fugu5.
In this page, two COPILOT datasets were merged into one Seurat object, with ALRA augmentation & CytoTRACE valuation.
for compatibility with COPILOT, a legacy version Seurat (v4.4) is applied.
```{r pkgs}
### Libraries
library(Seurat) #v4.4
library(ALRA)
library(CytoTRACE)

library(ggplot2)
library(pheatmap)

### Some usuful presets
ggcolor <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
`%notin%` <- Negate(`%in%`)

setwd("~/Library/CloudStorage/Box-Box/シングルセル解析(多部田)")
set.seed(101) #prevent the random output
```

## Merge two COPILOT datasets to one
```{r merge}
#(load two datasets)
wt <- readRDS("./1_scKB-COPILOT/WT_8DAS/WT_8DAS_COPILOT.rds")
fg5 <- readRDS("./1_scKB-COPILOT/fugu5_8DAS/fugu5_8DAS_COPILOT.rds")
wt
fg5

#(merge)
tbt1 <- merge(wt, y=fg5, add.cell.ids=c("WT","fugu5"), project="tbt_all")
Idents(tbt1) <- tbt1$orig.ident
tbt1

#(check the data integrity)
plot1 <- FeatureScatter(tbt1, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(tbt1, feature1 = "nCount_RNA", feature2 = "percent.cp")
plot3 <- FeatureScatter(tbt1, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2 + plot3
# subset outliners
tbt1 <- subset(tbt1, subset = percent.mt < 3 & percent.cp < 20 & nFeature_RNA < 3000)
#(re-normalize the count data by SCTransform)
tbt1 <- NormalizeData(tbt1, normalization.method = "LogNormalize", scale.factor = 10000)
tbt1 <- SCTransform(tbt1, variable.features.n = nrow(tbt1), assay = "RNA", new.assay.name = "SCT", verbose = FALSE) 
tbt1
```

## ALRA-augmentation & CytoTRACE
```{r alra & ct}
#(CytoTRACE)
results <- CytoTRACE(as.matrix(tbt1@assays$RNA@counts), ncores = 6, subsamplesize = 1000)
tbt1$CytoTRACE <- 1 - results$CytoTRACE

#(ALRA-imputation)
counts <- t(as.matrix(tbt1@assays$RNA@data)) #matrix where cells are rows and genes are columns
c_norm <- normalize_data(counts)
k_choice <- choose_k(c_norm) #38
#(details visualization for k)
require(gridExtra)
df <- data.frame(x=1:100,y=k_choice$d)
g1 <- ggplot(df,aes(x=x,y=y),) + geom_point(size=1)  + geom_line(size=0.5)+ geom_vline(xintercept=k_choice$k)   + theme( axis.title.x=element_blank() ) + scale_x_continuous(breaks=seq(10,100,10)) + ylab('s_i') + ggtitle('Singular values')
df <- data.frame(x=2:100,y=diff(k_choice$d))[3:99,]
g2 <- ggplot(df,aes(x=x,y=y),) + geom_point(size=1)  + geom_line(size=0.5)+ geom_vline(xintercept=k_choice$k+1)   + theme(axis.title.x=element_blank() ) + scale_x_continuous(breaks=seq(10,100,10)) + ylab('s_{i} - s_{i-1}') + ggtitle('Singular value spacings')
grid.arrange(g1,g2,nrow=1)

c_alra <- alra(c_norm, k=k_choice$k)[[3]]
# Read matrix with 7960 cells and 26512 genes
# Scaling all except for 2984 columns
# 0.00% of the values became negative in the scaling process and were set to zero
# The matrix went from 2.88% nonzero to 26.75% nonzero
c_alra.t <- as(t(c_alra), "dgCMatrix") 
colnames(c_alra.t) = rownames(counts)
tbt1[["ALRA"]] <- CreateAssayObject(data=c_alra.t)
```

## Save the progress
```{r save the progress}
saveRDS(tbt1, "Tabeta-fugu5.seurat.rds")
sessionInfo()
```

