---
title: "[Barley field transcriptome] Main Figure 1 and supplements"
author: "June-Sik Kim"
date: "2024-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Overview

R scripts to process the barley field transcriptome data and visualize them via Seurat. The raw and processed transcriptome data are available from NCBI GEO with a specific accession [GSE226906](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE226906 "Accession display").

Here is to load the required libraries and to setup the other environments.

```{r libraries}
library(Seurat) #ver.4.4.0
library(cluster) #for clusGAP
library(gplots) #for heatmap.2
library(glmnet) #for glm

#For GESA analysis
library(GSEABase) 
library(GOstats) #GO/KEGG
library(KEGGREST) #KEGG-API

#others
set.seed(101)
`%notin%` <- Negate(`%in%`) #temp. function to "not-in"

#emulate the ggplot hue() color palette
ggcolor <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
require(scales)
#key.palette <- hcl.colors(100, "Cividis")
#key.palette <- rev(hcl.colors(100, "RdYlBu"))
#show_col(key.palette)
show_col(ggcolor(8)) #test ggcolor
```

## Generate the Seurat object from RPM table
The RPM table is available from the NCBI GEO.
```{r seurat1}
rpm_f = "RPM_IPSR_KIBR_LB_17_19_BLD20201007.txt"
meta_f = "samplesheet.txt"

### Filter the metadata
meta <- read.table(meta_f, header=T)
rownames(meta) <- meta$SampleID
meta <- meta %>% filter(!Place %in% c("BLD"))
dim(meta) #[1] 1942   10

### Load the RPM table
rpm <- read.table(rpm_f, header=T, row.names = 1)
colnames(rpm) <- str_sub(colnames(rpm), 2) ## remove the first "X" from the colnames, using stringr()
rpm <- rpm[,meta$SampleID] ## filter only columns with metadata
dim(rpm) #[1] 39734  1942

### Create the Seurat object & supplement the metadata
so <- CreateSeuratObject(counts=rpm, project="so", min.cells=100, min.features=8000) #cut few samples by detected gene No.
meta <- meta[rownames(so@meta.data),]
so@meta.data$Place <- meta$Place
so@meta.data$Accession <- meta$Accession
so@meta.data$Year <- meta$Year
so@meta.data$Case <- meta$Case
so@meta.data$Date <- meta$Date
so@meta.data$Date.2 <- meta$Date.2
so@meta.data$Date.3 <- meta$Date.3
so@meta.data$fD2H <- meta$Date.4

### Refining the data & defining Highly Varying expressions
so = NormalizeData(so, scale.factor = 10^6) # log2RPM
so = ScaleData(so,do.scale = F, do.center = T) # Centering the normalized data to improve the performance of PCA, but not scaling, which might increase noise.
so <- FindVariableFeatures(so, selection.method = "mvp")
VariableFeatures <- VariableFeatures(so)
WholeFeatures <- rownames(so@assays$RNA$data)

so
# An object of class Seurat
# 25145 features across 1940 samples within 1 assay
# Active assay: RNA (25145 features, 2963 variable features)
```

## Figure 1A
```{r fig1A}
require(maps)
require(mapdata)
map(database="japan", xlim=c(125,150), ylim=c(30,50), resolution=0, interior=FALSE)
points(x=133.46, y=34.34, col=2, pch=19, cex=.5) #IPSR
points(y=35.44, x=139.638, col=2, pch=19, cex=.5) #KIBR
```

## PCA, ElasticNet-aided PC selection, UMAP
```{r seurat2}
# so <- RunPCA(so, vervose=FALSE, approx=FALSE, npcs=50, features=VariableFeatures)
# so <- RunUMAP(so, dims=c(1:50))
# so@reductions$pca_vari2 <- so@reductions$pca
# so@reductions$umap_vari2_50 <- so@reductions$umap
# saveRDS("seurat.barley-field.leaf.v3.rds")
so <- readRDS("seurat.barley-field.leaf.v3.rds")

### EN
alpha <- seq(0.01, 0.99, 0.01)
mse.df <- NULL
for (i in 1:length(alpha)) {
  m <- cv.glmnet(x = so@reductions$pca@cell.embeddings, y = so@meta.data$fD2H, family = "gaussian", alpha = alpha[i])
  mse.df <- rbind(mse.df, data.frame(alpha = alpha[i],
                                     mse = min(m$cvm)))
}
best.alpha <- mse.df$alpha[mse.df$mse == min(mse.df$mse)]
##[1] 0.96
#(cross-validation)
en.model.cv <- cv.glmnet(x = so@reductions$pca@cell.embeddings, y = so@meta.data$fD2H, family = "gaussian", alpha = best.alpha)
plot(en.model.cv, xvar="lambda", label=TRUE) 
#(EN-modeling)
en.model <- glmnet(x = so@reductions$pca@cell.embeddings, y = so@meta.data$fD2H, family = "gaussian", lambda = en.model.cv$lambda.1se)

abline(h=0.007,lty=2,col=2) #(set a threshold)
pcs <- which(abs(en.model$beta)>0.007) #(retrieve significant PCs)
#[1]  1  3 14 17 18 20 40
sum(sqrt(Stdev(so, reduction="pca")[pcs])) #(The sigma(beta))

so <- RunUMAP(so, dims=pcs)
so@reductions$umap_2D <- so@reductions$umap
```

## Figure 1E
```{r fig1E}
accs <- factor(so@meta.data$Accession, levels=c("J247","J064","J647","H602"))
boxplot(so@meta.data$nFeature_RNA ~ accs, col=ggcolor(4))
```

## Figure S2
```{r fig.s2}
## Figure S2A
VariableFeaturePlot(so)
## Figure S2B
ElbowPlot(so, ndims=50)
## Figure S2C
so@meta.data$Accession <- factor(so@meta.data$Accession, levels=c("J247","J064","J647","H602"))
DimPlot(so, reduction="umap_vari2_50", group.by="Accession")
## Figure S2D
FeaturePlot(so, reduction="umap_vari2_50", features="fD2H")
```

## Figure S3
```{r figs3}
## Figure S3A
plot(en.model.cv, xvar="lambda", label=TRUE)
## Figure S3B
plot(abs(en.model$beta), pch=19, cex=0.3)
abline(h=0.007,lty=2,col="grey80") 
```

## Figure 1F-H
```{r fig1F-H}
## Figure 1F
FeaturePlot(so, reduction="umap_2D", features="fD2H")
## Figure 1G
so@meta.data$month <- factor(so@meta.data$month, levels=c(12,1,2,3,4,5))
DimPlot(so, reduction="umap_2D", group.by="month", cols=rev(ggcolor(6))[order(levels(so@meta.data$month))])
## Figure 1H
so@meta.data$Accession <- factor(so@meta.data$Accession, levels=c("J247","J064","J647","H602"))
DimPlot(so, reduction="umap_2D", group.by="Accession", cols=ggcolor(4)[c(3,2,1,4)])
```

## Figure S4
```{r figs4}
## Figure S4A
key.palette <- rev(hcl.colors(100, "Lajolla"))
FeaturePlot(so, reduction="umap_2D", features="DayHour", col=key.palette)
## Figure S4B
key.palette <- hcl.colors(100, "Inferno")
FeaturePlot(so, reduction="umap_2D", features="AvgTemp", col=key.palette)
## Figure S4C
DimPlot(so, reduction="umap_2D", group.by="Batch", cols=hcl.colors(4, "Fall"))
```

## Session Info
```{r sessioninfo}
sessionInfo()
```
